---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


#This is for my notes only - please look at PA1_template.md for the assignment


```{r firstplot, results="markup"}
with(data, 
     plot(date, steps, 
          xlab="date",
          ylab="steps",
          main = "Steps By Day",
          type="h", lwd=4))
```





```{r}
cdata <- na.omit(data)  #remove NAs
origMean <- as.data.frame(tapply(cdata$steps, cdata$date, mean)) #calculate mean, by date
setDT(origMean, keep.rownames = TRUE)[]
colnames(origMean) <- c("date", "mean_of_steps_byDate")

```




```{r}
origMedian <- as.data.frame(tapply(cdata$steps, cdata$date, median))
setDT(origMedian, keep.rownames = TRUE)[]
colnames(origMedian) <- c("date", "median_of_steps_byDate")
```


```{r}
with(data2, 
     plot(date, steps, 
          xlab="date",
          ylab="steps",
          main = "Steps By Day, Na=Mean by Interval",
          type="h", lwd=4))
```


```{r}
with(data2, 
     plot(date, steps, 
          xlab="date",
          ylab="steps",
          main = "Steps By Day, Na=Mean by Interval",
          type="h", lwd=4))
```


```{r}
newMean <- as.data.frame(tapply(data2$steps, data2$date, mean))

```
```{r}
newMedian <- as.data.frame(tapply(data2$steps, data2$date, median))

```




par(mfrow=c(2,1))
with(intervalWK, plot(namesDatwk, wkMeans, 
                       xlab="Time Intervals", 
                       ylab="Average Steps, over all days",
                       main= "Weekdays only", 
                       type="l", 
                       lwd=4, 
                       col="blue"))
with(intervalWKEnd, plot(namesDatwkEnd, wkEndMeans, 
                       xlab="Time Intervals", 
                       ylab="Average Steps, over all days",
                       main= "Weekend only", 
                       type="l", 
                       lwd=4, 
                       col="red"))
x<- c(0,500,1000,1500,2000,2500)
axis(side=1, at=x, labels=T)
#axis(1, at = seq(10, 2000, by = 1000))


## Loading and preprocessing the data



## What is mean total number of steps taken per day?



## What is the average daily activity pattern?



## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?


###Weekday Only Plot Creation
```{r }
weekOnly <- data2[which(data2$wDay == 'weekday'), ]
wkMeans <- tapply(weekOnly$steps, weekOnly$interval, mean) #creates array of intervals and their means
wkintDat <- as.data.frame(wkMeans)

namesDatwk <- names(wkMeans) #takes names of array, sets to variable #must be a faster way to do this
intervalWK <- cbind(namesDatwk, wkintDat)
```

###Weekends Only Plot Creation
```{r}
endOnly <- data2[which(data2$wDay == 'weekend'), ]
wkEndMeans <- tapply(endOnly$steps, endOnly$interval, mean) #creates array of intervals and their means
wkEndintDat <- as.data.frame(wkEndMeans)
namesDatwkEnd <- names(wkEndMeans) #takes names of array, sets to variable #must be a faster way to do this
intervalWKEnd <- cbind(namesDatwkEnd, wkEndintDat)
```

### Weekend & Weekday Plots
```{r fourth plot}
par(mfrow=c(2,1))
with(intervalWK, plot(namesDatwk, wkMeans, 
                       xlab="Time Intervals", 
                       ylab="Average Steps, over all days",
                       main= "Weekdays only", 
                       type="l", 
                       lwd=4, 
                       col="blue"))
with(intervalWKEnd, plot(namesDatwkEnd, wkEndMeans, 
                       xlab="Time Intervals", 
                       ylab="Average Steps, over all days",
                       main= "Weekend only", 
                       type="l", 
                       lwd=4, 
                       col="red"))
#axis(1, at = seq(10, 2000, by = 1000))
```
